# Introduction to Landscape Ecology {#week1}

## Requirements

**Chapter 1** - Introduction to Landscape Ecology. [@gergel_2001]

## Learning objetives

* Install R and R Studio in your computer
* Understand the basic datatypes and data structures in R
* Declare basic datatypes and data structures
* Do basic element-wise and vector-wise operations

## Lab Theory

### What is R?

* R is a multiplatform programming language (*"It means that we ask the computer to do jobs."*)
* R works in Windows, MacOs, Linux.
* R is **free**.
* Easy to learn and very powerful.
* R can do many jobs: 
    + Used as a calculator.
    + R can be used to create figures such as scatter plots, bar plots, pie plots, etc.
    + R can be used to summarize data (numerically and graphically).
    + R can be used to do a lot of specific analyses.
    
### What is R Studio?

* RStudio is an integrated development environment (IDE) for R. 
* It includes a console, syntax-highlighting editor that supports direct code execution. 
* Tools for plotting, history, debugging and workspace management.
* Open source.

### Installation of R and R Studio

* In this course, we run R on Windows for demonstration purposes.
* The installation of R on Windows is easy:
    + Download R (for Windows) at https://cran.r-project.org.
    + Run the downloaded .exe file.
    + Current version is 3.5.2 (upgraded multiple times every year).
* Installation of RStudio is also straightforward
    + Download can be found https://www.rstudio.com/products/rstudio/download/#download
    + Current version 1.1.463 (for windows).

## Lab: Getting familiar with R

Installation

<iframe width="560" height="315" src="https://www.youtube.com/embed/ttshzOISCsE" frameborder="0" allow="accelerometer; autoplay; encrypted-media; gyroscope; picture-in-picture" allowfullscreen></iframe>

Calculation and variables

<iframe width="560" height="315" src="https://www.youtube.com/embed/S-o-sdlzhkE" frameborder="0" allow="accelerometer; autoplay; encrypted-media; gyroscope; picture-in-picture" allowfullscreen></iframe>

Create and Work With Vectors

<iframe width="560" height="315" src="https://www.youtube.com/embed/YhQOV27pQfg" frameborder="0" allow="accelerometer; autoplay; encrypted-media; gyroscope; picture-in-picture" allowfullscreen></iframe>

Character and Boolean Vectors

<iframe width="560" height="315" src="https://www.youtube.com/embed/GKu5tw_bIpA" frameborder="0" allow="accelerometer; autoplay; encrypted-media; gyroscope; picture-in-picture" allowfullscreen></iframe>

Vector Arithmetic

<iframe width="560" height="315" src="https://www.youtube.com/embed/u5hroyx0J4o" frameborder="0" allow="accelerometer; autoplay; encrypted-media; gyroscope; picture-in-picture" allowfullscreen></iframe>

## Lab: Code practice

This section is intended to complement the concepts explained in videos. You will find all the commands used but also the expected results. Please type these codes in your RStudio Editor and start to execute (CTRL+ENTER) line by line.

### Task 1: Basic variables:

```{r, results='markup'}
numberStudents <- 41          # Declaration of an integer value
myFirtsName <- "Abraham"      # Declaration of a character value
myWeight <- 160.5             # Declaration of a numeric value
isUCStudent <- TRUE           # Declaration of a logical value
height <- c(152, 171.5, 165)  # Declaration of a numeric vector
mat <- matrix(    
              c(14, 16, 18, 15, 22, 11),
              nrow = 3,
              ncol = 2,
              byrow = TRUE
              )
# Declaration of a n-dimensional matrix
rubikCube <- array(c('green','yellow','red','blue', 'white','orange'),dim = c(3,3,6)) 
#Declaration of a data frame
friends <- data.frame( names = c("Thomas","Andrew","Lindsay"), gender = c("Male", "Male","Female"), height = c(152, 171.5, 165), weight = c(81,93, 78), Age = c(42,38,26) )

# See the content of your variables
print(numberStudents)         # Check the content of numberStudents
print(myFirtsName)            # Check the content of myFirstName
print(myWeight)               # Check the content of myWeight
isUCStudent                   # Additional way to see the content within
dim(mat)                      # Check the dimensions of the matrix
print(mat)                    # Check the content of the mat matrix 
dim(rubikCube)                # Check the dimensions of the matrix
print(rubikCube)              # Check your rubik content
print(friends)                # Check list content
```

### Task 2: Operations 
```{r}
a <- 1.5                      # Declaration of a numeric value
b <- 4.0                      # Declaration of another numeric value

a + b                         # Summation
b - a                         # Substraction
b * a                         # Multiplication
b / a                         # Division
b ** 2                        # Exponent
b %% 2                        # Modulus operation

# Declaration of a vector of numeric values
vector_values <- c(1, 4, 9, 12, 15, 23, 17, 13, 9) 

mean(vector_values)           # Computing the average of the set of values.
sum(vector_values)            # summation of the set of values.
length(vector_values)         # Computing the average of the set of values.
vector_values + 3             # Add or substract a number to each value on the set of values.
vector_values * 5             # Multiply a number to each value on the set of values.
vector_values %% 2            # Apply a modulo operation over the set of values.
vector_values[1]              # Indexing vectors in R (find the ith element).
vector_values[5]              # Suppose that we want to find the fifth element.
vector_values[-2]             # Get the vector without the second element.
vector_values[1:3]            # Get the first three elements of the vector.
vector_values[c(1,3,5)]       # Get the elements 1st, 3rd and 5th of the vector.
vector_values == 9            # Check if the vector values are equals to 9
which(vector_values == 9)     # where are this value
sum(vector_values == 9)       # How many 9's are in the vector
```

### Task 3. Creating vectors
```{r}
new_vector <- 1:10            # Create a vector from 1 to 10
new_vector
new_vector <- rep(5, 10)      # Create a vector of 10 5's (ten times five)
new_vector
new_vector <- seq(1, 10, 2)   # Create a vector from 1 to less equals to 10 but with step of 2
new_vector
new_vector <- seq(2, 10, 2)   # Same as before but with even numbers
new_vector
```

### Task 4. Operations between two or more vectors
```{r}
vector1 <- seq(2,10,2) 
vector1
vector2 <- seq(11,20,2)       # Declare two vectors
vector2
vector_result <- c(vector1, vector2) # Create a vector with the combination of vector1 and vector2
vector_result
vector1 + vector2             # Addition of two vectors
vector1 - vector2             # Substraction of two vectors
vector1 * vector2             # Multiplication of two vectors
vector1 / vector2             # division of two vectors
max(vector_result)            # vector max
min(vector_result)            # vector min
pmax(vector1, vector2)        # element-wise max
pmin(vector1, vector2)        # element-wise min
head(vector_result)           # first 5 elements
```

## Project 1.
* Create a vector *x* with 100 elements, the first 50 are 1, 2, 3, ... , 99 and the rest 50 elements are 2, 4, ..., 100.
* Create a vector *y* with 100 elements, the first 50 element are all 100; the rest 50 elements are 1, 2, 3, ..., 50.
* Calculate *x+y* and the element-wise max for *x* and *y*; the sum of elements in *x* and *y*; the maximum value of *x*.

## Project 2. 
Create a variable
```{r}
x <- c(2, 9, 3, 7, 4, 9, 6, 6, 6, 6)
```
Compute the following: 

* (x1 + x2, ..., x10)/10.
* Find Log10(xi) for each *i*.
* Find (xi - 1.5)/2.6 for each *i*.
* Find the difference between the largest and smallest values of *x*. (This is the range. You can use *max* and *min*.)

## Project 3. 
Suppose you keep track of one variable *x* at some soecific poiny in the time. The last 8 values were:

55311 35624 15908 36219 65499 67251 65745 34447

* Enter this numbers in R
    + Use commands diff on the data. What does it give?
    + Find the maximum number of the variable, the average number and the minimum number between the sample times.  

## Project 4. 
Suppose one of your bills varies from month to month. Also suppose your year has the following monthly
amounts:

50 23 42 39 47 30 31 21 12 15 55 58

* Enter this data into R using a variable called bill.
* Find the amount you spent this year on your bills.
* What is the smallest amount you spent?
* What is the largest?
* How many bills was the amount greater than $40?
* What percentage of the total bills was this?