  # Dataframes, matrices and lists in R {#week2}

Here is a review of existing methods.

## Recap

* Vector operations and generation
* Logical operations:

```{r echo=TRUE, results='hide'}
  ">" # for "greater than"
  ">=" # for "greater than or equal to"
  "<" # for "less than"
  "<=" # for "less than or equal to"
  "==" # for "equal to"
  "!=" # for "not equal to"
  "&" #for "and"
  "|" # for "or"
```

* Create vectors:
```{r echo=TRUE, results='asis'}
  new_vector <- 1:10 # Create a vector from 1 to 10
  new_vector
  new_vector <- rep(5, 10) # Create a vector of 10 5's (ten times five)
  new_vector
  new_vector <- seq(1, 10, 2) # Create a vector from 1 to less equals to 10 but with step of 2
```

## Introduction to Dataframes

* Usually the data is stored in a tabular form, with rows for different person and columns for different variables.
* Suppose that we have the following table:

```{r echo = TRUE, results='asis'}
plot_id <- c(3, 7, 4, 4, 7, 8, 2, 3, 7, 8, 8, 1, 7, 4, 4, 6, 19, 23, 18)
species_id <- c("DM","DM","DM","DM","DM","DO","PF","OX","DM","PF","DM","PF","DM","PF","DM","DM","DM","DM","DM")
sex <- c('M','M','F','F','M','F','M','F','F','M','F','M','F','F','F','F','F','F','F')
hindfoot_length <- c(35,37,34,35,35,32,15,21,36,12,32,16,34,14,35,37,35,35,33)
weight <- c(40,48,29,46,36,52,8,22,35,7,22,9,42,8,41,37,43,41,40)
study <- data.frame(plot_id,species_id,sex,hindfoot_length,weight)
```

* We can add row names to the data frame
```{r echo = FALSE, results='asis'}
rownames(study) <- 1:19
```

* or change column names
```{r echo = TRUE, results='asis'}
colnames(study)<-c("ID", "NAME", "SEX", "HINDFOOT_LENGHT", "WEIGHT") # Assign names to the dataframe column
```

* Now we can summarize a column,
```{r echo = TRUE, results='asis'}
mean(study[,4]) # Compute the mean from a numeric vector
sd(study[,4]) # Compute the standard error from a vector
```

* Notice that we select a column by the following command:
```{r echo = TRUE, results='asis'}
(study)[,4]
```

* Question: how to select a row?
```{r echo = TRUE, results='asis'}
study[4,]
```

* We can also select a column or a row by its name
```{r echo = TRUE, results='asis'}
study[,"NAME"] # Call column by name
```

To select a entry, we can do
```{r echo = TRUE, results='asis'}
study[1,2]
```

this is because that a table is two dimensional.

* Question: how to select the first and second column and make it a new data frame?

* Question: how to select the records (rows) for all females and make it a new data frame?

## Lists

* Lists are the R objects which contain elements of different types like (numbers, strings, vectors) and another list inside it. A list can also contain a matrix or a function as its elements. List is created using list() function.

* Let us create a list: 
```{r echo = TRUE, results='asis'}
list.example <- list( a = 1:5, b = c("Cincinnati", "Columbus", "Cleveland"), c = study)
```

* What is in the list?
* To access a list element, we can use the "$" operator:

```{r echo = TRUE, results='asis'}
list.example$"a"
```

Or the following:
```{r echo = TRUE, results='asis'}
list.example[["a"]]
list.example[[1]]
```

* What if you use single square brackets "[" instead of "[["?
```{r echo = TRUE, results='asis'}
list.example[2] # Access the second element of the list but not the elements within
typeof(list.example[2]) # Type of element with single brackets (a list with the elements of the first position)
typeof(list.example[[2]]) # Type of element with double brackets
```

* A data frame is a list, where all elements are the column vectors.
So we can access a column of a data frame using "$" operator:
```{r echo = TRUE, results='asis'}
  study$"SEX"
```

* Try the following commands:
```{r echo = TRUE, results='asis'}
  length(list.example) # get the length of the list 
  length(list.example[1]) # get the length of the list returned with the elements of the first position (some ackward notation of r) 
  length(list.example[[1]]) # get the length of the element in the first position of the list 
```

## Matrices

A matrix is an extension of vector. It has additional dimension attributes (row and columns).

* Let us create a matrix:
```{r echo = TRUE, results='asis'}
matrix.example <- matrix(1:9, nrow = 3) # Create a matrix with 9 elements and 3 rows (columns are automatically adjusted).
matrix.example
matrix.example.2 <- matrix(1:8, nrow = 2) # Create a matrix with 8 elements (1 to 8) and 2 rows.
matrix.example.2
```

* We can change a matrix to a vector (ignore the dimension information for each entry):

```{r echo = TRUE, results='asis'}
as.vector(matrix.example.2) # get a vector (Dimension 1) with the elements of matrix.example.2
```

* How to know the dimmension of a matrix:
```{r echo = TRUE, results='asis'}
dim(matrix.example.2) # Dimension of matrix.example.2
```

* What if we want to create a matrix with the elements sorted by row 
```{r echo = TRUE, results='asis'}
matrix.example.byrow <- matrix( 1:9, nrow = 3, byrow = TRUE) # Create a matrix with elements by row. 
matrix.example.byrow
```
* matrices also support operations
```{r echo = TRUE, results='asis'}
matrix.example + matrix.example.byrow # Adding two matrices
matrix.example - matrix.example.byrow # Substracting two matrices
```

* The same logic to access the elements
```{r echo = TRUE, results='asis'}
matrix.example[1,] # Accessing row 1
matrix.example[,2] # Accessing row 2
matrix.example[1,2] # Accessing row 1,2
```
