---
output:
  word_document: default
  pdf_document: default
  html_document: default
---
# Mapping in R {#week6}

## Basics

### Tiles

_"A tile map (raster or vector) is a map displayed in a browser by seamlessly joining dozens of individually requested image files over the internet."_

Let's create a simple map:

```{r}
library(leaflet)   # load the library
mymap<-leaflet() %>%
  addTiles # images you are going to see
mymap   #check the map
```

Now, it is time to explore different tiles:

CartoDB:

```{r}
mymap<-leaflet() %>%
  addProviderTiles(provider ="CartoDB.Positron")
mymap   #check the map
```

ESRI World Imagery:

```{r}
mymap<-leaflet() %>%
  addProviderTiles(provider ="Esri.WorldImagery")
mymap   #check the map
```

Night lights:

```{r}
mymap<-leaflet() %>%
  addProviderTiles(provider ="NASAGIBS.ViirsEarthAtNight2012")
mymap   #check the map
```

Now that we have the basic map, let's populate it with some fun projects!

## Project 1: Load data points

Download the datasets available in canopy (*csv):

```{r,eval=FALSE}
data<-read.csv("MOCK_DATA.csv")

mymap<-leaflet(data = data) %>%
  addTiles %>% # images you are going to see
  addMarkers(lng = ~long,
             lat = ~lat,
             popup = ~city)
mymap   #check the map
```

* Remember these dataset have no real data, it was created for learning purposes.

## Project 2: Processing your data

```{r, eval=FALSE}
data<-read.csv("MOCK_DATA2.csv")

mymap<-leaflet(data=data) %>%
  addTiles %>% # images you are going to see
  addMarkers(lng = ~long,
             lat = ~lat,
             popup = ~country)
mymap   #check the map

# EX 3  Filter by US
rm(list = ls())
data<-read.csv("MOCK_DATA2.csv")

onlyUS<-data[data$country=="US",]

cleanUS<-na.omit(onlyUS)

mymap<-leaflet(data = cleanUS) %>%
  addTiles %>% # images you are going to see
  addMarkers(lng = ~long,
             lat = ~lat,
             popup = ~income)
mymap   #check the map
```

## Project 3: Adding a legend

```{r, eval=FALSE}
rm(list = ls())
data<-read.csv("MOCK_DATA2.csv")

onlyUS<-data[data$country=="US",]

cleanUS<-na.omit(onlyUS)

pal = colorNumeric("RdYlBu", domain = cleanUS$income)

mymap<-leaflet(data = cleanUS) %>%
  addTiles %>% # images you are going to see
  addCircles(lng = ~long,
             lat = ~lat,
             popup = ~city,
             radius = ~sqrt(income)*200,
             color = ~pal(income),
             fillOpacity = 1,
             opacity = 1) %>%
  addLegend(title = "Income",pal = pal, values = ~income,position = "bottomright") 
mymap   #Generate your map
```

### Homework

1. Using the fake data depicts North America cities (Mexico, Canada, United States).

2. Create your own dataset with the top 10 largest cities in Ohio. Remember to include for each city:
+ City name
+ lat and long coordinates 
+ Income
+ Population
+ The expected answer should be a CSV file

3. Load your Ohio dataset into the leaflet map and show the city as pop up message.

4. Add the proper circles and legends. (Bonus point)
  
### Useful links

1. https://mockaroo.com/

2. https://www.latlong.net/

3. https://rstudio.github.io/leaflet/
